---
title: "London Modelling Notebook"
author: "You!"
output:
  html_document: 
    keep_md: yes
---



## Set up covered so far in the slides
Load up packages and the data
```{r}
library(tidyverse)
library(sf)
library(INLA)
library(spdep)
## effects and gridExtra used in slides but not really required

## only need to do this once, can also do it manually from the GitHub page
if(!any(grepl('LondonMSOA',list.files()))){
  download.file('https://github.com/smithtr/SpatialHealthOGH/raw/main/Main_Session/Data/LondonMSOA.zip',
              'LondonMSOA.zip')
  unzip('LondonMSOA.zip')
}

MSOA_London = st_read('./LondonMSOA/London_MSOAs_Vaccines.shp')
London_Vax = select(MSOA_London, vax_rate, n, IMD,young, old, geometry)%>%
  mutate(Y = round(vax_rate/100*n, digits = 0))

head(London_Vax)

```

Basic GLM
```{r}
logit_base = glm(cbind(Y, n-Y)~ IMD+ I(IMD^2)+ old + young*I(young<0.4),
                 data = London_Vax, family = 'binomial')
```


Making the adjacency matrix for London
```{r}
London_NB = poly2nb(MSOA_London)
London_W = nb2mat(London_NB, style = 'B')
```


Add in some columns for random effects index:
```{r}
London_Vax_INLA = mutate(London_Vax, IDv = 1:nrow(London_Vax), IDu = IDv)
```


## Modelling with R-INLA
### Fit the INLA model

I've given you the general structure for the call. Fill in the RHS of the formula
```{r, eval = FALSE}
London_BYM = inla(Y ~ ..., 
                  Ntrials = n, data = London_Vax_INLA, family = 'binomial')
```

### Posterior summaries
Various ways to look at the output of INLA

```{r, eval = FALSE}
names(INLA)
plot(London_BYM)
London_BYM$summary.fixed
London_BYM$summary.hyperpar
London_BYM$summary.random
```

Plot of spatial random effect
```{r, eval = FALSE}
London_Vax$u = London_BYM$summary.random$IDu$mean
ggplot() + 
  geom_sf(data = London_Vax, aes(fill = u))
```

### Exceedence

Question: which MSOAs have substantially lower odds of vaccination after accounting for
deprivation and age structure?


Answer: Find areas where the following quantity is high

$$p(v_i+u_i < \text{threshold})$$


```{r}
London_BYMv2 = inla(Y ~ IMD+ I(IMD^2)+ old + young*I(young<0.4) + f(IDu, model = "bym", graph = London_W),
                    Ntrials = n, data = London_Vax_INLA, family = 'binomial')

?inla.pmarginal
```

Choose your own cutoff
```{r, eval = FALSE}
exceed_probs = vector(length = nrow(London_Vax))
for(i in 1:nrow(London_Vax)) exceed_probs[i] = inla.pmarginal(q = ...., marginal = London_BYMv2$marginals.random$ID[[i]])

London_Vax$exceed_probs =  exceed_probs

ggplot() + 
  geom_sf(data = London_Vax, aes(fill = exceed_probs))

```



### Package versions
```{r}
sapply(sessionInfo()$otherPkgs, function(x){x$Version})
```


