---
title: "Spatial Health Analysis Challenge Code"
output:
  html_document: 
    keep_md: yes
---

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(sf)
```


Read in the polygon (you will need to download from the data folder first).
```{r, message=FALSE}
map_MSOA = st_read('./Data/Middle_Layer_Super_Output_Areas_2011/Middle_Layer_Super_Output_Areas_2011.shp')
MSOA_bath = map_MSOA[grep('Bath', map_MSOA$msoa11nm),]
```

Download the case data for Bath from UK Covid dashboard API.
First MSOA-level data:  rolling seven day totals, released once per week.
Missing rows mean censored data (counts = 0, 1, or 2).
 
```{r}
cases_bath_msoa = read_csv('https://api.coronavirus.data.gov.uk/v2/data?areaType=msoa&areaCode=E06000022&metric=newCasesBySpecimenDateRollingSum&format=csv')
```

Next get data for the whole local authority: rolling seven day totals, released daily.
```{r}
cases_bath_all = read_csv('https://api.coronavirus.data.gov.uk/v2/data?areaType=utla&areaCode=E06000022&metric=newCasesBySpecimenDateRollingSum&format=csv')
```
Note: The UK covid-19 API is sometimes down. These above data sets are also available 
in the data folder on GitHub. 
  
  * weely_cases_bath_MSOA contains the MSOA-level rolling seven day totals, released every week.
  * daily_cases_bath_total contains the rolling seven day totals for the whole local authority, released daily

 
Now I extract data for one low week between peaks in April 2021 (lots of censoring)
and one peak week in July 2021 to demonstrate the censoring problem.
```{r, message=FALSE}
low_week = filter(cases_bath_msoa, as.character(date) == '2021-04-20')%>%
  select(areaCode, April_2021 = contains('Cases'))


high_week = filter(cases_bath_msoa, as.character(date) == '2021-07-27')%>%
  select(areaCode, July_2021 = contains('Cases'))

low_week
high_week
```  



Next check the total counts for these weeks, and save some labels for plots later.
```{r, message=FALSE}
filter(cases_bath_all, as.character(date) == '2021-04-20' | as.character(date) ==  '2021-07-27')
wk.labs <- c("20 April 2021 (B&NES total 16)", "27 July 2021 (B&NES total 634)")
names(wk.labs) <- c("April_2021", "July_2021")
```

Now I merge with MSOA geometry and reshape the data structure a bit to make plotting easier
```{r, message=FALSE}
MSOA_bath = left_join(MSOA_bath, low_week, by = c('msoa11cd' = 'areaCode'))%>%
  left_join(high_week,by = c('msoa11cd' = 'areaCode'))

msoab2 = select(MSOA_bath, April_2021, July_2021, geometry) %>% 
  gather(week, cases, -geometry)
```

Finally I make an unfancy plot, which shows the general problem 

```{r, message=FALSE}
ggplot() + 
  geom_sf(data = msoab2, aes(fill = cases)) + 
  scale_fill_continuous(low="thistle2", high="darkred",
                        guide="colorbar",na.value="gray95")+
  facet_wrap(~week, ncol = 2, labeller = labeller(week = wk.labs))+
  theme_bw()+ ggtitle('Confirmed Cases in Bath and Northeast Somerset')
```
